---
title: "Abandoned_Props_GeoCoding"
author: "Sharif Nijim, East Section"
date: "December 5, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# initialize the environment
rm(list=ls()) 
# load the necessary libraries
setwd("C:/Users/Owner/Desktop/Data Science/Data Viz/Data-Viz-2018-Fall-master/FinalProject")
library(lubridate)
library(reshape2)
library(tidyverse)
library(rgdal)
library(sp)
library(tmap)
library(ggmap)
register_google(key = "AIzaSyArtVYaFH3qvcVP2ufW5_LMMANBwZK2vS4")
library(leaflet)
options(scipen=999)
```

This week, your assignment has three parts:

First, you will load pre-existing spatial data into R and create a quick static map using spplot.  
Second, you will geocode a small subset of a larger dataset.  
Finally, you will create a leaflet object from any of these datasets or one of the other examples I have provided you.  
# Part A: Using R with spatial data (7 points)  
School Boundaries  
## 1. Load spatial data from a shapefile into R (3 points)  
### read in the data
```{r}
# read in the data
# Provide the code you used to load the file  
ap <- readOGR(dsn="Abandoned_Property_Parcels", layer = "Abandoned_Property_Parcels", stringsAsFactors = FALSE)
apdf <- as.data.frame(ap@data)
```
```{r}
# display a summary of the data
# Provide the output of the object summary  
summary(ap)
```
```{r}
# Show what is in the data
data.frame(ap)
```
### Create a static map showing this layer
```{r}
# Create a static map showing this layer
qtm(ap)
```

```{r}
ap.spatial <- SpatialPointsDataFrame(coords = ap@data[,c("Lon","Lat")],
                                       data = ap@data,
                                       proj4string = CRS("+proj=longlat +datum=WGS84"))
```
```{r, eval=FALSE}
# geocode?
#3622 EDISON RD, SOUTH BEND, IN 46615
#ap@data$Address_Nu + ap@data$Street_Nam + ap@data$Zip_Code
ap.num <- as.character(ap@data$Address_Nu)
ap.dir <- as.character(ap@data$Direction)
ap.street <- as.character(ap@data$Street_Nam)
ap.suff <- as.character(ap@data$Suffix)
ap.zip <- as.character(ap@data$Zip_Code)
ap.city <- "South Bend, IN"
ap.outcome <- melt(as.character(ap@data$Outcome_St))
#numstreet <- rbind(ad.num, ad.street)
thelist <- vector()
for (i in 1:length(ap.num)) {
  thelist[[i]] <- paste(ap.num[[i]],
          na.omit(ap.dir[[i]]),
          ap.street[[i]],
          ap.suff[[i]],
          ", South Bend, IN",
          ap.zip[[i]])
}
ap.full.addrs <- melt(thelist)
ap.geo <- ap.full.addrs
```

```{r, eval=FALSE}
# GEOCODE
# loop through and get the lat/long from the address
#for (i in 1:nrow(ap.full.addrs)) {
for (i in 1:nrow(ap.full.addrs)) {
  tp <- geocode(as.character(ap.full.addrs$value[i]),
                output = "latlona",
                source = "google")
  ap.geo$lon[i] <- as.numeric(tp[1])
  ap.geo$lat[i] <- as.numeric(tp[2])
  ap.geo$accuracy[i] <- as.character(tp[3])
}
# Fail: 807 N Olive ST , South Bend, IN 46628, line 1212
for (i in 1213:nrow(ap.full.addrs)) {
  tp <- geocode(as.character(ap.full.addrs$value[i]),
                output = "latlona",
                source = "google")
  ap.geo$lon[i] <- as.numeric(tp[1])
  ap.geo$lat[i] <- as.numeric(tp[2])
  ap.geo$accuracy[i] <- as.character(tp[3])
}
```
```{r}
apdf$lat <- ap.geo$lat
apdf$lon <- ap.geo$lon
apdf$accuracy <- ap.geo$accuracy
```
```{r, eval=FALSE}
# save the data
save(ap.num,
     ap.dir,
     ap.street,
     ap.suff,
     ap.zip,
     ap.city,
     thelist,
     ap.full.addrs,
     ap.geo,
     apdf,
     file = "Abandoned_Properties_Geocoded.v4.Rdata")
```